---
title: "H2020: Exploratory data analysis of auxiliary data"
author: "Jakob Etzel"
date: "26 Feb 2018"
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lubridate)
library(rworldmap)
library(knitr)
library(randomcoloR)
library(rje)
newmap <- getMap(resolution = "low")
source(file = "/home/datascience/h2020gridforecast/R/hdf5_data_layer.R")
```

## GOSAT
Format/extension: h5.
Measurements available for 5 consecutive days, but each with varying (amounts) of geolocations.
Contents details see the [products page](http://www.gosat.nies.go.jp/en/about_5_products.html), section L2 data.

[Greenhouse gases Observing SATellite (GOSAT) Data Policy](https://data2.gosat.nies.go.jp/doc/GOSAT_Data_PolicyB_en.pdf)

| file name | product code | product name |
|---|---|---|
| gosat_FTS_C01S_2 | C01S | L2 CO2 column amount (SWIR) |
| gosat_FTS_C02S_2 | C02S | L2 CH4 column amount (SWIR) |
| gosat_FTS_C03S_2 | C03S | L2 H2O column amount (SWIR) |

```{r data retrieval, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
gosat <- get_gosat()
gosat$H2O$day = date(ymd_hms(gosat$H2O$timestamp))
gosat$CO2$day = date(ymd_hms(gosat$CO2$timestamp))
gosat$CH4$day = date(ymd_hms(gosat$CH4$timestamp))
days <- gosat$H2O %>% select(day) %>% unique() %>% pull(day)
```

We have data from the following days: `r days`.

### Relation between the three products
Overall, the timeseries are measuring exactly simultaneously (both time and geolocation). However, there are two data points in C02S which are not present in C01S (time, lon, lat):
```{r}
(gosat$CO2 %>% select(timestamp, lon, lat))[c(449, 697),]
```

Between C01S and C03S the first 448 entries are synchronous. From then on, longitudinal values differ completely, i.e. no longitudinal value of either the C01S or C02S time series occurs anywhere in the C03S time series. This is because the second data batch is not of days "2010-07-13" and "2010-07-14", but "2010-07-18" and "2010-07-19".

### World Map Plot

We plot all the measurement geolocations and colour them according to the measurement day.

```{r echo=FALSE}
selectInput(
  inputId = "gosat_product", 
  label = "GOSAT product:", 
  choices=names(gosat)
)
```

```{r global map, echo=FALSE}
renderPlot({
  plot(newmap, asp = 1)
  product_geolocations <- gosat[[input$gosat_product]] %>% select(lon, lat, day)
  days <- product_geolocations$day %>% unique()
  palette <- distinctColorPalette(k = length(days))
  for (i in 1:length(days)) {
    points(product_geolocations %>% filter(day == days[i]) %>% pull(lon), product_geolocations %>% filter(day == days[i]) %>% pull(lat), col = palette[i], cex = .6)
  }
})
# points((gosat$CO2 %>% filter(day == days[1]))$lon, (gosat$CO2 %>% filter(day == days[1]))$lat, col = "red", cex = .6)
# points((gosat$CO2 %>% filter(day == days[2]))$lon, (gosat$CO2 %>% filter(day == days[2]))$lat, col = "blue", cex = .6)
# points((gosat$CO2 %>% filter(day == days[3]))$lon, (gosat$CO2 %>% filter(day == days[3]))$lat, col = "green", cex = .6)
# points((gosat$CO2 %>% filter(day == days[4]))$lon, (gosat$CO2 %>% filter(day == days[4]))$lat, col = "orange", cex = .6)
# points((gosat$CO2 %>% filter(day == days[5]))$lon, (gosat$CO2 %>% filter(day == days[5]))$lat, col = "yellow", cex = .6)
```

Now same data zoomed in on Europe:

```{r Europe map, echo=FALSE}
renderPlot({
  plot(newmap, xlim = c(-20, 59), ylim = c(35, 71), asp = 1)
  product_geolocations <- gosat[[input$gosat_product]] %>% select(lon, lat, day)
  days <- product_geolocations$day %>% unique()
  palette <- distinctColorPalette(k = length(days))
  for (i in 1:length(days)) {
    points(product_geolocations %>% filter(day == days[i]) %>% pull(lon), product_geolocations %>% filter(day == days[i]) %>% pull(lat), col = palette[i], cex = .6)
  }
})
```

### CO2 data
Out of a lot of available columns I tried to select the most relevant ones. *Timestamp* is the time of the satellite measurement, *day* is the date of the timestamp, *lat* and *lon* define the area of the measurement (somehow the centre of the inspection area), *dryAirColumn* and *surfacePressure* are auxiliary data measured; *CO2TotalColumn* and *xco2* (mixing ratio) are the main measurements. *surfaceAlbedo*, *surfaceWindSpeed* and *temperatureProfile* are reference data from an external source.

As an example, here the first five entries. Afterwards the histograms of the measurements (over all days and ovservations).

```{r sample data, echo=FALSE}
kable(x = gosat$CO2[1:5,])
hist(gosat$CO2$CO2TotalColumn)
hist(gosat$CO2$XCO2)
hist(gosat$CO2$dryAirTotalColumn)
hist(gosat$CO2$surfacePressure)
hist(gosat$CO2$surfaceAlbedo)
hist(gosat$CO2$surfaceWindSpeed)
hist(gosat$CO2$temperatureProfile)
```

It seems as if for *albedo* and *temperature* the value -9999 equals a non-existing or faulty value. We therefore plot again the histgrams ommitting these data points:

```{r sample data extra hists, echo=FALSE}
hist(gosat$CO2 %>% filter(temperatureProfile != -9999) %>% pull(temperatureProfile))
hist(gosat$CO2 %>% filter(surfaceAlbedo != -9999) %>% pull(surfaceAlbedo))
```

## Open Power Systems Data (OPSD) {.tabset}
Format/extension: h5.
Prepared by competition organisers from downloaded .csv files (15 min and 60 min interval data), it has been converted to .h5 format and sliced to be in sync with the target data by the competition organisers.

[About Open Power System Data](https://open-power-system-data.org/about/)
[README file](https://data.open-power-system-data.org/time_series/2017-07-09/README.md)

Data includes *([15] means 15 min resolution; others in 60 min resolution) -- only those columns listed where time series is non-empty*:

* capacity
    + solar (DE) [15]
    + wind offshore/onshore/total (DE [50Hz, Amprion, Tennet, total]) [15]
* generation
    + wind offshore/onshore (AT, DE [50HZ, Amprion, Tennet, TransnetBW, total], DK [1, 2], SE)
    + solar (DE [Amprion, Tennet], SE)
* forecast
    + wind (DE [50Hz, Amprion, Tennet, TransnetBW])
    + solar (DE [50Hz, Amprion, Tennet])
* wind profile *Percentage of wind capacity producing in Germany* (DE)
* load (AT, BA, BE, BG, CH, CY, CZ, DE, DK, EE, ES, FI, FR, GB [total, NIR], GR, HR, HU, IE, IS, IT, , LT, LU, LC, ME, MK, NL, NO, PL, PT, RO, RS, SE, SI, SK)
* price day-ahead (DE, DK [1, 2], NO, SE)

### OPSD 15 (train)
Time resolution of the data is 15 minutes. We list here the column names and the average measurement values for each column. The frequent occurences of NaN as an average value show that most of the time series are completely empty.
```{r, echo=FALSE}
aux_opsd_15_train <- readRDS(file = "../../../local_data/aux_opsd_15_train.rds")
data.frame("column name" = colnames(aux_opsd_15_train), "average value" = unlist(lapply(colnames(aux_opsd_15_train), function (x) mean(aux_opsd_15_train %>% pull(x), na.rm = TRUE))))

aux_opsd_15_train_not_NaN <- readRDS(file = "../../../local_data/aux_opsd_15_train_not_NaN.rds")
aux_opsd_15_train_not_NaN$datetime = as.POSIXct(aux_opsd_15_train_not_NaN$timestamp, origin = "1970-01-01")
```

The start and end of the data are:
```{r}
summary(aux_opsd_15_train_not_NaN %>% select(datetime))[1]
summary(aux_opsd_15_train_not_NaN %>% select(datetime))[6]
```

We plot now the non-empty time series:
```{r echo=FALSE}
selectInput(
  inputId = "opsd_15_train_columns", 
  label = "OPSD columns:", 
  choices=colnames(aux_opsd_15_train_not_NaN)
)
```

```{r echo=FALSE}
renderPlot({
  plot(sort(aux_opsd_15_train_not_NaN$datetime), aux_opsd_15_train_not_NaN %>% arrange (datetime) %>% pull(input$opsd_15_train_columns), type = "l")
  title(main = input$opsd_15_train_columns)
})
```

### OPSD 15 (adapt)
Available in the starter kit's `auxiliaryData/adapt` folder. Time resolution of the data is 15 minutes. Data is structured in a lot of different files which are numbered and should mimic forecasting steps, i.e. data updates being received every hour. Therefore, the single files only contain one timestep each.

For these hourly measurements, we have 57 different columns available.
We list here the column names and the average measurement values for each column. The frequent occurences of NaN as an average value show that most of the time series are completely empty.

```{r, echo=FALSE}
aux_opsd_15_adapt <- readRDS(file = "../../../local_data/aux_opsd_15_adapt.rds")
data.frame("column name" = colnames(aux_opsd_15_adapt), "average value" = unlist(lapply(colnames(aux_opsd_15_adapt), function (x) mean(aux_opsd_15_adapt %>% pull(x), na.rm = TRUE))))

aux_opsd_15_adapt_not_NaN <- readRDS(file = "../../../local_data/aux_opsd_15_adapt_not_NaN.rds")
aux_opsd_15_adapt_not_NaN$datetime = as.POSIXct(aux_opsd_15_adapt_not_NaN$timestamp, origin = "1970-01-01")
```

The start and end of the data are:
```{r}
summary(aux_opsd_15_adapt_not_NaN %>% select(datetime))[1]
summary(aux_opsd_15_adapt_not_NaN %>% select(datetime))[6]
```

We plot now the non-empty time series:

```{r echo=FALSE}
selectInput(
  inputId = "opsd_15_adapt_columns", 
  label = "OPSD columns:", 
  choices=colnames(aux_opsd_15_adapt_not_NaN)
)
```

```{r echo=FALSE}
renderPlot({
  plot(sort(aux_opsd_15_adapt_not_NaN$datetime), aux_opsd_15_adapt_not_NaN %>% arrange(datetime) %>% pull(input$opsd_15_adapt_columns), type = "l")
  title(main = input$opsd_15_adapt_columns)
})
```

### OPSD 60 (adapt)
Available in the starter kit's `auxiliaryData/adapt` folder. Time resolution of the data is 60 minutes. Data is structured in a lot of different files which are numbered and should mimic forecasting steps, i.e. data updates being received every hour. Therefore, the single files only contain one timestep each.

The numbering of the files shows steps - there is no no. 48, 49. Also the order of the timestamps seems not to be monotone.

```{r}
list.dirs(
  path = "/home/datascience/starting-kit/auxiliaryData/adapt/aux/opsd60/", 
  full.names = FALSE
)[-1]
```

For these hourly measurements, we have 259 different columns available.
We list here the column names and the average measurement values for each column. The frequent occurences of NaN as an average value show that most of the time series are completely empty.

```{r, echo=FALSE}
aux_opsd_60_adapt <- readRDS(file = "../../../local_data/aux_opsd_60_adapt.rds")
data.frame("column name" = colnames(aux_opsd_60_adapt), "average value" = unlist(lapply(colnames(aux_opsd_60_adapt), function (x) mean(aux_opsd_60_adapt %>% pull(x), na.rm = TRUE))))

aux_opsd_60_adapt_not_NaN <- readRDS(file = "../../../local_data/aux_opsd_60_adapt_not_NaN.rds")
aux_opsd_60_adapt_not_NaN$datetime = as.POSIXct(aux_opsd_60_adapt_not_NaN$timestamp, origin = "1970-01-01")
```

The start and end of the data are:
```{r}
summary(aux_opsd_60_adapt_not_NaN %>% select(datetime))[1]
summary(aux_opsd_60_adapt_not_NaN %>% select(datetime))[6]
```

We plot now the non-empty time series:
```{r echo=FALSE}
selectInput(
  inputId = "opsd_60_adapt_columns", 
  label = "OPSD columns:", 
  choices=colnames(aux_opsd_60_adapt_not_NaN)
)
```

```{r echo=FALSE}
renderPlot({
  plot(sort(aux_opsd_60_adapt_not_NaN$datetime), aux_opsd_60_adapt_not_NaN %>% arrange(datetime) %>% pull(input$opsd_60_adapt_columns), type = "l")
  title(main = input$opsd_60_adapt_columns)
})
```

### Wind/solar forecast
There are non-empty time series available for wind forecasts for the four German transmission system operators (TSOs) TenneT, 50Hertz, Amprion, TransnetBW. Additionally, there are solar forecasts available for TenneT, Amprion.

We examine these time series based on the [50 Hertz (TSO) wind data archive](http://www.50hertz.com/de/Kennzahlen/Windenergie/Archiv-Windenergie) of the year 2010. Predictions are available there in a 15 minute resolution (sic!). We can read the following values:

```{r echo=FALSE}
data.frame(
  date = c("09.07.10", "09.07.10", "09.07.10", "09.07.10"),
  start = c("05:00", "05:15", "05:30", "05:45"),
  end = c("05:15", "05:30", "05:45", "06:00"),
  forecast = c("926,38", "888,46", "850,55", "812,11")
)
```

The average forecast value for this given hour is 896,375 which exactly ressembles the value of the respective aux opsd 60 time series:
```{r}
(aux_opsd_60_adapt_not_NaN %>% select(datetime, DE_50hertz_wind_forecast))[1,]
```

These values are [provided by 50 Hertz in advance](http://www.50hertz.com/de/Kennzahlen/Windenergie/Prognose) in the way that a quarter hour X on day Y in the data means the forecast for X and Y and it is published at day Y-1 18:00 for the public.

On the contrary, in the auxiliary data provided by the competition organisers, values for quarter hour X on day Y are simulated to be made public only on day Y at the last full hour, this ranges from X-1 to X-4. Thus we loose most of its predictive character.

### Comparison
The OPSD 15 train and adapt data set have the same columns. All OPSD 15 columns also appear in OPSD 60. For OPSD 15, the numbers of non-empty columns differ between training and adapt data set.
```{r}
min(colnames(aux_opsd_15_adapt) == colnames(aux_opsd_15_train))
is.subset(colnames(aux_opsd_15_adapt), colnames(aux_opsd_60_adapt))
length(colnames(aux_opsd_15_adapt_not_NaN)) == length(colnames(aux_opsd_15_train_not_NaN))
```

## NOAA {.tabset}

### HYCOM
So far no analysis done.

### NCOM
So far no analysis done.