---
title: "Impact of auxiliary variables"
author: "Jakob Etzel"
date: "26 Feb 2018"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(h2020gridforecast)
library(dplyr)
source(file = "../config.R")

train_all <- readRDS(file = "../../../local_data/train_all.rds")
aux_opsd_15_train_not_NaN <- readRDS(file = '../../../local_data/aux_opsd_15_train_not_NaN.rds')
```

## Impact of selected OPSD columns
We have a look at the target train data and the OPSD 15 aux train data. We measure correlations between every third target measurement and every OPSD 15 measurement (15 min = 3 * 5 min). For a given OPSD 15 time series, we calculate correlations on the full available time span for every power line. We then give quantiles and a histogram for these correlation coefficients.

For some variables, we see that there are a couple of power lines with a correlation coefficient above 0.5:
* DE_tennet_wind_offshore_generation
* 

We could use the OPSD time series as an exogeneous variable to explain those power lines.

```{r, echo=FALSE, warning=FALSE}
selectInput(
  inputId = "opsd15_column", 
  label = "OPSD 15 time series:", 
  choices=colnames(aux_opsd_15_train_not_NaN %>% select(-timestamp))
)
cors <- reactive({
  unlist(lapply(1:1916, function (x) cor(aux_opsd_15_train_not_NaN %>% pull(input$opsd15_column), (train_all %>% pull(x))[0:105048 %% 3 == 0][1:11750])))
})
renderPrint({
  summary(cors())
})
renderPlot({
  hist(cors())
})
```

